---
title: "Simple Linear Regression"
subtitle: "Introduction"
author: "<br><br> Dr. Maria Tackett"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: "slides.css"
    logo: "img/introregression-sticker.png"
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
      slideNumberFormat: ""
    seal: false #make custom title slide
editor_options: 
  chunk_output_type: console
---

```{r child = "setup.Rmd"}
```

class: title-slide 

<br><br>

# Simple Linear Regression 
## Introduction 

<br><br><br>

### Dr. Maria Tackett

---

## Topics 

- Use simple linear regression to describe the relationship between a quantitative predictor and quantitative response variable.

- Estimate the slope and intercept of the regression line using the least squares method.

- Interpret the slope and intercept of the regression line.

```{r packages, echo = F}
library(tidyverse)
library(broom)
library(knitr)
library(kableExtra)
library(fivethirtyeight) #fandango dataset
```

```{r data}
movie_scores <- fandango %>%
  rename(critics = rottentomatoes, 
         audience = rottentomatoes_user)
```

???

I want to see how the presenter notes work

---

## Movie ratings data 

The data set contains the "Tomatometer" score (**`critics`**) and audience score (**`audience`**) for 146 movies rated on rottentomatoes.com.

```{r}
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  labs(x = "Critics Score" , 
       y = "Audience Score") +
  theme_bw()
```

??? 

Do the presenter notes work?? Whoo! 
---

## Movie ratings data 

We want to fit a line to describe the relationship between the critics score and audience score.

```{r}
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(x = "Critics Score" , 
       y = "Audience Score") +
  theme_bw()
```

---

## Terminology

.pull-left[
The .vocab[response, *Y*], is the variable describing the outcome of interest.

<br> 

The .vocab[predictor, *X*], is the variable we use to help understand the variability in the response.
]

.pull-right[
```{r, fig.height = 4, }
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(x = "Critics Score (Predictor)" , 
       y = "Audience Score (Response)") +
  theme_bw() +
  theme(axis.title=element_text(size=20))
```
]

---

## Regression model 

A .emph[regression model] is a function that describes the relationship between the response, $Y$, and the predictor, $X$. 

.eq[
$$\begin{aligned} Y &= \color{black}{\textbf{Model}} + \text{Error} \\[8pt]
&= \color{black}{\mathbf{f(X)}} + \epsilon \\[8pt]
&= \color{black}{\boldsymbol{\mu_{Y|X}}} + \epsilon \end{aligned}$$
]

---

class: middle, center

.pull-left[
.small-box[
$$\begin{aligned} Y &= \color{purple}{\textbf{Model}} + \text{Error} \\[8pt]
&= \color{purple}{\mathbf{f(X)}} + \epsilon \\[8pt]
&= \color{purple}{\boldsymbol{\mu_{Y|X}}} + \epsilon \end{aligned}$$
]
]

.pull-right[
```{r, fig.height = 4}
m <- lm(audience ~ critics,data=movie_scores)
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "purple", se = FALSE) +
  labs(x = "X" , 
       y = "Y") +
  theme_bw() +
  theme(axis.text = element_blank(),
  axis.ticks.x = element_blank(), 
  axis.ticks.y = element_blank()
  ) 
```
]

---

class: middle

.pull-left[
.small-box[
$$\begin{aligned} Y &= \color{purple}{\textbf{Model}} + \color{blue}{\textbf{Error}} \\[5pt]
&= \color{purple}{\mathbf{f(X)}} + \color{blue}{\boldsymbol{\epsilon}} \\[5pt]
&= \color{purple}{\boldsymbol{\mu_{Y|X}}} + \color{blue}{\boldsymbol{\epsilon}} \\[5pt]
 \end{aligned}$$
]
]

.pull-right[
```{r, fig.height = 4}
m <- lm(audience ~ critics,data=movie_scores)
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "purple", se = FALSE) +
  geom_segment(aes(x=critics, xend=critics, y=audience, yend=predict(m)), color = "blue") +
  labs(x = "X" , 
       y = "Y") +
  theme_bw() +
  theme(axis.text = element_blank(),
  axis.ticks.x = element_blank(), 
  axis.ticks.y = element_blank()
  ) 
```
]

---

## Simple linear regression

When we have a quantitative response, $Y$, and a single quantitative predictor, $X$, we can use a .vocab[simple linear regression] model to describe the mean of $Y$ given $X$.
.eq[
$$\begin{aligned} Y &= \mathbf{\beta_0 + \beta_1 X} + \epsilon \end{aligned}$$
]

$$\boldsymbol{\beta}_1: \text{Slope} \hspace{20mm} \boldsymbol{\beta}_0: \text{Intercept}$$

---

class: middle, center

.eq[
$$\Large{\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 X}$$
]



---

class: middle

.eq[
$$\Large{\hat{Y} = \hat{\beta}_0 + \hat{\beta}_1 X}$$
]

---

class: middle, center

How do we choose values for $\hat{\beta}_1$ and $\hat{\beta}_0$?

```{r fig.height = 2.5}
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.4) + 
  geom_abline(intercept = 32.296, slope = 0.5187, color = "red", size = 1) +
  geom_abline(intercept = 25, slope = 0.7, color = "steel blue") +
  geom_abline(intercept = 21, slope = 0.9, color = "steel blue") +
  geom_abline(intercept = 35, slope = 0.3, color = "steel blue") +
  labs(x = "Critics Score" , 
       y = "Audience Score") +
  theme_bw()
```



---

## Residuals

```{r echo=F, warning=FALSE, message=FALSE, fig.height = 2.5}
m <- lm(audience ~ critics,data=movie_scores)
ggplot(data=movie_scores,mapping=aes(x=critics,y=audience)) + 
  geom_point(alpha=0.5) + 
  geom_smooth(method="lm", color = "red", se=FALSE) + 
  geom_segment(aes(x=critics, xend=critics, y=audience, yend=predict(m)), color = "steel blue") +
  labs(x = "Critics Score" , 
       y = "Audience Score")+
  theme_bw() + 
  theme(legend.position = "none")
```

.eq[
$$\text{residual} = \text{observed} - \text{predicted} = y - \hat{y}$$ 
]

---

## Least squares line

- The residual for the $i^{th}$ observation is

$$e_i = \text{observed} - \text{predicted}
= y_i - \hat{y}_i$$
--

- The .vocab[sum of squared residuals] is 

$$e^2_1 + e^2_2 + \dots + e^2_n$$

--

- The .vocab[least squares line] is the one that minimizes the sum of squared residuals

---

```{r}
sx <- round(sqrt(var(movie_scores$critics)),4)
sy <- round(sqrt(var(movie_scores$audience)),4)
r <- round(cor(movie_scores$critics, movie_scores$audience),4)
xbar <- round(mean(movie_scores$critics),4)
ybar <- round(mean(movie_scores$audience),4)
```

## Estimating the slope
.eq[
$$\large{\hat{\beta}_1 = r \frac{s_Y}{s_X}}$$
]

--

.pull-slight-left[
.small-box-work[
$$\begin{aligned} &s_X = 30.169 \\[5pt]
&s_Y =  20.024 \\[5pt]
&r = 0.781 \end{aligned}$$
]
]

--

.pull-right[
.small-box-work[
$$\begin{aligned}\hat{\beta}_1 &= 0.781 \times \frac{20.024}{30.169} \\[10pt]
&= \mathbf{0.518}\end{aligned}$$
]
]

---

## Estimating the intercept

.eq[

$$\large{\hat{\beta}_0 = \bar{Y} - \hat{\beta}_1\bar{X}}$$
]

--

.pull-slight-left[
.small-box-work[
$$\begin{aligned}&\bar{x} = 60.850 \\[5pt]
&\bar{y} = 63.877 \\[5pt]
&\hat{\beta}_1 = 0.518 \end{aligned}$$
]
]

--

.pull-more-right[
.small-box-work[
$$\begin{aligned}\hat{\beta}_0 &= 63.877 - 0.518 \times 60.850 \\[10pt]
&= \mathbf{32.296}\end{aligned}$$
]
]

---

## Interpreting slope & intercept

.eq[
$$\hat{\text{audience}} = 32.296 + 0.518 \times \text{critics}$$
]

<br>

--

.vocab[Slope]: For each additional point in the critics score, we expect the audience score to increase by 0.518 points, on average.

--

.vocab[Intercept]: If the critics score is 0, we expect the audience score to be 32.296 points. 

---

## Does it make sense to interpret the intercept?

`r emo::ji("white_check_mark")` **Interpret the intercept if**
  - the predictor can feasibly take values equal to or near zero.
  - there are values near zero in the data.

<br> 
`r emo::ji("stop_sign")` Otherwise, don't interpret the intercept!

---

## Recap

- Use simple linear regression to describe the relationship between a quantitative predictor and quantitative response variable.

- Use the least squares method to estimate the slope and intercept.

- **Interpretations**:
  - **Slope**: For each additional value of $x$, we expect y to change by $\hat{\beta}_1$ units, on average. 
  - **Intercept**: If $x$ is 0, then we expect $y$ to be $\hat{\beta}_0$ units.
 

