## SLR Notes 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T)
```

This is text to see what is happening. 

```{r pressure}
plot(pressure)
```

I would like to embed slides into the docs. 

<iframe seamless width="672" height="400px" frameborde="0" src="../slides/03-slr.html"></iframe>



This from slides that need to go in the notes: 

- It was originally used in the article [*Be Suspicious of Online Movie Ratings, Especially Fandango's*](https://fivethirtyeight.com/features/fandango-movies-ratings/) by FiveThirtyEight.

- Variables: 
    - **`critics`**: Tomatometer score for the film (0 - 100)
    - **`audience`**: Audience score for the film (0 - 100)

```{r}
glimpse(movie_scores)
```


small[
```{r}
p1 <- ggplot(data = movie_scores,mapping = aes(x = critics)) +
  geom_histogram()
p2 <- ggplot(data = movie_scores,mapping = aes(x = audience)) +
  geom_histogram()
p1 + p2
```


audience is the response variable (Y)
variable whose variation we want to understand and/or variable we wish to predict
also known as dependent, outcome, target variable


critics is the predictor variable (X)
variable used to account for variation in the response
also known as independent


<br><br>
*Optional: [Deriving the Least-Squares Estimates for Simple Linear Regression](https://github.com/sta210-sp20/supplemental-notes/blob/master/slr-derivations.pdf)*

