## IN-CLASS: Movies on IMDB

```{r echo = F}
knitr::opts_chunk$set(eval = F, 
                      warning = F,
                      message = F
                      )
```

In this activity, we will examine the relationship between budget and revenue for movies made in the United States in 1986 to 2016. The data is originally [@imdb].

```{r warning=FALSE, message=FALSE}
#load packages
library(readr)
library(tidyverse)
library(DT)
```

### Data {-}

The dataset `movies.csv` includes basic information about each movie including budget, genre, movie studio, director, etc. A full list of the variables may be found [here](https://github.com/danielgrijalva/movie-stats).

```{r read-data, message=FALSE, warning=FALSE}
movies <- read_csv("https://raw.githubusercontent.com/danielgrijalva/movie-stats/master/movies.csv")
```

```{r filter-data}

movies <- movies %>%
  filter(country=="USA", #use movies in the USA
         !(genre %in% c("Musical","War","Western"))) #removes genres with < 10 movies
```

```{r view-data}
#view the data
movies
```

### Analysis {-}

#### Part 1 {-}

We begin by observing how the gross revenue (`gross`) has changed over time. Since we want to visualize the results, we will choose a few genres of interest for the analysis. 

```{r country-list}
genre_list <- c("Horror", "Drama", "Action", "Animation")
```

```{r plot-rating, fig.width=10, warning=FALSE}
movies %>%
  filter(genre %in% genre_list) %>% 
  group_by(genre,year) %>%
  summarise(avg_gross = mean(gross)) %>%
  ggplot(mapping = aes(x = year, y = avg_gross, color=genre)) +
    geom_line() +
    ylab("Average Gross Revenue (in US Dollars)") +
    ggtitle("Gross Revenue Over Time")
```

#### Part 2 {-}

Next, let's see the relationship between a movie's budget and its gross revenue. Because there is a large range of values for budget and revenue, we will plot the log-transformed version of each variable to more easily visualize the relationship. We will talk more about variable transformations later in the semester.

```{r plot-revenue, fig.width=10, warning=FALSE}
movies %>%
  filter(genre %in% genre_list, budget > 0) %>% 
  ggplot(mapping = aes(x=log(budget), y = log(gross), color=genre)) +
  geom_point() +
  geom_smooth(method="lm",se=FALSE) + 
  xlab("Log-transformed Budget")+
  ylab("Log-transformed Gross Revenue") +
  facet_wrap(~ genre)
```

### Next Steps {-}

1. Put your name in the author field in the `YAML` at the top of the .Rmd file. Knit again. (We will talk more about the `YAML` later in future assingments.)

2. Change the genre names in parts 1 and 2 to genres that interest you. The spelling and capitalization must match what's in the data, so you can use the [Appendix](#appendix) to see the correct spelling and capitalization. Knit again.

### Discussion Questions {-}

1. Consider the plot in Part 1. 
    - Describe how movie revenue has changed over time. 
    - Suppose we use revenue as a measure of popularity. How has the popularity of each genre changed over time? In other words, are the genres that were most popular in 1986 still the most popular today?

2. Consider the plot in Part 2. 
    - Which genre(s) tend to have the highest budgets?
    - In general, what is the relationship between a movie's budget and its total revenue? Are there any genres that show a different relationship between budget and revenue?

### Appendix {-}

Below is a list of genres in the data set:

```{r}
movies %>% 
  arrange(genre) %>% 
  select(genre) %>%
  distinct() %>%
  datatable()
```

***