## IN-CLASS: Advertising & Sales

```{r echo = F}
knitr::opts_chunk$set(
  eval = F
)
```

In this mini analysis, we will work with the `Advertising` data used in Chapters 2 and 3 of [@islr].


### Packages {-}


```{r message=FALSE}
library(readr)
library(tidyverse)
library(skimr)
library(broom)
```

### Data {-}
```{r}
advertising <- read_csv("data/advertising.csv")
```

This dataset contains the advertising spending and sales for 200 markets. The variables are

- `tv`: total spending on TV advertising (in $thousands)
- `radio`: total spending on radio advertising (in $thousands)
- `newspaper`: total spending on newspaper advertising (in $thousands)
- `sales`: total sales (in $millions)

### Exercises {-}

We'll begin the analysis by getting quick view of the data: 

```{r}
glimpse(advertising)
```

Next, we can calculate summary statistics for each of the variables in the data set. 

```{r}
advertising %>% skim()
```

1. What type of advertising has the smallest median spending?

2. What type of advertising has the largest variation in spending? 

3. Describe the shape of the distribution of `sales`. 

We are most interested in understanding how advertising spending affect sales. One way to quantify the relationship between the variables is by calculating the correlation matrix. 

```{r}
advertising %>% 
  cor()
```

1. What is the correlation between `radio` and `sales`? Interpret this value. 

2. What type of advertising has the strongest linear relationship with `sales`? 

Below are visualizations of `sales` versus each explanatory variable. 

```{r}
advertising %>%
  ggplot(mapping = aes(x=tv,y=sales)) + 
  geom_point(alpha=0.7) + 
  geom_smooth(method="lm",se=FALSE,color="blue") + 
  labs(title = "Sales vs. TV Advertising", 
       x= "TV Advertising (in $thousands)", 
       y="_____") #fill in the Y axis label
```

```{r}
advertising %>%
  ggplot(mapping = aes(x=radio,y=sales)) + 
  geom_point(alpha=0.7) + 
  geom_smooth(method="lm",se=FALSE,color="red") + 
  labs(title = "Sales vs. TV Advertising", 
       x= "Radio Advertising (in $thousands)", 
       y="Sales (in $millions)")
```

```{r}
advertising %>%
  ggplot(mapping = aes(x=newspaper,y=sales)) + 
  geom_point(alpha=0.7) + 
  geom_smooth(method="lm",se=FALSE,color="purple") + 
  labs(title = "Sales vs. Newspaper Advertising", 
       x= "Newspaper Advertising (in $thousands)", 
       y="Sales (in $millions)")
```

Since `tv` appears to have the strongest linear relationship with `sales`, let's calculate a simple linear regression model using these two variables. 

```{r}
ad_model <- lm(sales ~ tv, data=advertising)
ad_model
```

1. Write the model equation. 

2. Does it make sense to interpret the intercept? If so, interpret the intercept in the context of the problem. Otherwise, briefly explain why not.

3. Interpret the slope in the context of the problem.

***
