## IN-CLASS: Advertising & Sales

```{r echo = F}
knitr::opts_chunk$set(
  eval = F
)
```

In this mini analysis, we will work with the `Advertising` data used in Chapters 2 and 3 of *Introduction to Statistical Learning*.

### Data and packages

We start with loading the packages we'll use.

```{r message=FALSE}
library(readr)
library(tidyverse)
library(skimr)
library(broom)
```

```{r}
advertising <- read_csv("data/advertising.csv")
```

We will analyze the advertising and sales data for 200 markets. The variables we'll use are 

- `tv`: total spending on TV advertising (in $thousands)
- `radio`: total spending on radio advertising (in $thousands)
- `newspaper`: total spending on newspaper advertising (in $thousands)
- `sales`: total sales (in $millions)

### Analysis

We'll begin the analysis by getting quick view of the data: 

```{r}
glimpse(advertising)
```

Next, we can calculate summary statistics for each of the variables in the data set. 

```{r}
advertising %>% skim()
```

1. What type of advertising has the smallest median spending?

2. What type of advertising has the largest variation in spending? 

3. Describe the shape of the distribution of `sales`. 

We are most interested in understanding how advertising spending affect sales. One way to quantify the relationship between the variables is by calculating the correlation matrix. 

```{r}
advertising %>% 
  cor()
```

1. What is the correlation between `radio` and `sales`? Interpret this value. 

2. What type of advertising has the strongest linear relationship with `sales`? 

Below are visualizations of `sales` versus each explanatory variable. 

```{r}
advertising %>%
  ggplot(mapping = aes(x=tv,y=sales)) + 
  geom_point(alpha=0.7) + 
  geom_smooth(method="lm",se=FALSE,color="blue") + 
  labs(title = "Sales vs. TV Advertising", 
       x= "TV Advertising (in $thousands)", 
       y="_____") #fill in the Y axis label
```

```{r}
advertising %>%
  ggplot(mapping = aes(x=radio,y=sales)) + 
  geom_point(alpha=0.7) + 
  geom_smooth(method="lm",se=FALSE,color="red") + 
  labs(title = "Sales vs. TV Advertising", 
       x= "Radio Advertising (in $thousands)", 
       y="Sales (in $millions)")
```

```{r}
advertising %>%
  ggplot(mapping = aes(x=newspaper,y=sales)) + 
  geom_point(alpha=0.7) + 
  geom_smooth(method="lm",se=FALSE,color="purple") + 
  labs(title = "Sales vs. Newspaper Advertising", 
       x= "Newspaper Advertising (in $thousands)", 
       y="Sales (in $millions)")
```

Since `tv` appears to have the strongest linear relationship with `sales`, let's calculate a simple linear regression model using these two variables. 

```{r}
ad_model <- lm(sales ~ tv, data=advertising)
ad_model
```

1. Write the model equation. 

2. Interpret the intercept in the context of the problem. 

3. Interpret the slope in the context of the problem.

***
