## IN-CLASS: Carbohydrates in Beer

```{r echo = F}
knitr::opts_chunk$set(eval = FALSE, 
                      warning = FALSE,
                      message = FALSE
                      )
```

### Packages {-}
```{r}
library(tidyverse)
library(readr)
library(broom)
```


### Data {-}
```{r}
beer <- read_csv("data/beer.csv")
```

In this analysis, we will analyze the relationship between the amount of alcohol (`PercentAlcohol`) and the caloric content (`CaloriesPer12Oz`) in domestic beers. Let `PercentAlcohol` be the predictor variable and `CaloriesPer12Oz` the response variable. 

### Exercises {-}

**Due to limited class time, we will not do the exploratory data analysis in this example. In practice, however, you should always start with the exploratory data analysis.**


1. Fit a regression model to describe the relationship between `PercentAlcohol` and `CaloriesPer12Oz`. Display the model output.

```{r}
# code for regression model
```

2. Does it make sense to interpret the intercept? Why or why not?

3. Interpret the 95% confidence interval for the slope in the context of the data.

4. Calculate the critical value, $t^*$, used to calculate the 95% confidence interval. The code below is a guide; uncomment and complete the lines of code to calculate and display the critical value.

```{r}
n <- nrow(beer)

#df <- _________  

#crit_val <- qt(_____,df)
#crit_val
```

The critical value used to calculate the 95% confident interval for the slope is ____. 

5. Interpret the test statistic in the context of the data.

6. How was the p-value calculated? Fill in the code below to calculate the p-value. The code below is a guide; uncomment and complete the lines of code to calculate and display the p-value.

```{r}
#test_statistic <- _______

#prob <- 1 - pt(abs(test_statistic),df)

#p_value <- 2 * prob
#p_value
```

The p-value is ____ . Given there is no linear relationship between `PercentAlcohol` and `CaloriesPer12Oz`, the probability of obtaining a test statistic with magnitude ____ or more extreme is ____.

7. Fill in the code below to calculate the predicted calories and corresponding 90% interval for a single beer with alcohol content of 4.3%.

```{r}
x0 <-  data.frame(PercentAlcohol=4.3)
#predict.lm(model,x0,interval="________",conf.level=________) #<<
```


8. Fill in the code below to calculate the predicted calories and corresponding 90% interval for the subset of beers with alcohol content of 4.3%.

```{r}
x0 <-  data.frame(PercentAlcohol=4.3)
#predict.lm(model,x0,interval="________",conf.level=________) #<<
```

***