## IN-CLASS: Diamonds - A Review Exercise

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                   message=FALSE, 
                   eval = FALSE)
```

### Packages {-}
```{r}
library(tidyverse)
library(broom)
library(rms)
library(knitr)
```

### Data {-}
```{r}
set.seed(12)
diamonds_samp <- ggplot2::diamonds %>% 
  filter(carat < 1.1) %>%
  sample_n(300) %>%
  mutate(log_price = log(price), 
         caratCent = carat - mean(carat), 
         caratCent_sq = caratCent^2, 
        color = factor(as.character(color)), # to fix variable format in model output
        clarity = factor(as.character(clarity)) # to fix variable format in model output
)
```

### Exercises {-}

#### Main Effects Model {-}

- Why should we use `logprice` instead of `price` as the response variable? In other words, what is an example of previous analysis that could have been done to help us determine whether to use `logprice` or `price`?

Below is the model with `log_price` as the response and `caratCent`, `color`, and `clarity` as the predictor variables.

```{r}
model_orig <- lm(log_price ~ caratCent +  color + clarity, data=diamonds_samp)
kable(tidy(model_orig),format="markdown")
```

- What is the baseline level of `color`? What is the baseline level of `clarity`?

- Interpret the intercept in terms of `price`.

```{r}
coef <-  model_orig$coefficients
```

- Calculate and describe the difference in the typical prices of diamonds that are color E and diamonds that are color D, holding all else constant.

```{r}
## calculate difference
```


- Calculate and describe the difference in the typical prices of diamonds that are color E and diamonds that are color G, holding all else constant.

```{r}
## calculate difference
```

- What is the predicted price of a single diamond that has color E, clarity VS2 and is 0.3 carats? Finish the code below the predicted value and the corresponding interval.

```{r}
x0 <- data.frame(color="___", clarity="___", carat=___)
x0 <- x0 %>% mutate(
  caratCent = carat - mean(____),
  caratCent_sq = caratCent^2
)

(exp(predict(model_orig,x0,interval=____))) #interval to predict for single observation
```


- Suppose we wish to find the predicted median price of subset of all diamonds with color E, clarity VS2, and 0.3 carats. How do you expect the predicted price to change? How do you expect the corresponding interval to change?

-  Write code to find the predicted price and corresponding interval for the median price for the subset of all diamonds with color E, clarity VS2 and 0.3 carats.

```{r}
# code for predicted price
```
<br><br>

Use the code below to obtain the ANOVA table for this model.

```{r}
anova(model_orig)
```

- What is the estimated regression variance?

- Compare $R^2$ and $Adj. R^2$. What does this comparison tell you about the predictors in the model?

- Use the code below to calculate the VIF for this model.

```{r}
vif(model_orig)
```

- This model has potential problems with multicollinearity. How did we come to this conclusion? Which variables are highly collinear?

- Why do you think this multicollinearity is occurring? *Hint: Examine the distribution of the variable(s) that have high multicollinearity.*

#### Model with Interactions {-}

- Below is the model that includes an interaction term between `caratCent` and `color`. What is the appropriate method to determine if the interaction is significant? 

```{r}
model_int <- lm(log_price ~ caratCent + color + 
                  clarity + caratCent*color, data=diamonds_samp)
```

- Conduct the test you listed above. What is your conclusion? 

```{r}
# code for test
```

*Regardless of your answer to the previous question, use `model_int` to answer the next two questions:*


- What is the estimated slope of `caratCent` for a diamond with `color==D`?

- What is the estimated slope of `caratCent` for a diamond with `color==J`?


***
