## Multinomial Logistic Regression 

```{r echo = F}
knitr::opts_chunk$set(eval = FALSE, 
                      warning = FALSE,
                      message = FALSE
                      )
```


```{r,echo=F}
library(tidyverse)
library(knitr)
library(broom)
library(cowplot)
library(arm) #binned residuals
library(nnet) #multinomial regression models
```

The main objective of this analysis is to understand how encouragement affects the frequency that children watch *Sesame Street*. We will use the following variables: 

**Response:**

- <font class="vocab">`viewcat`</font>
    + 1: rarely watched show
    + 2: once or twice a week
      + 3: three to five times a week
      + 4: watched show on average more than five times a week


**Predictors:**

- <font class="vocab">`age`:</font> child's age in months
- <font class="vocab">`prenumb`: </font>score on numbers pretest (0 to 54)
- <font class="vocab">`prelet`: </font>score on letters pretest (0 to 58)
- <font class="vocab">`viewenc`:</font> 1: encouraged to watch, 2: not encouraged
- <font class="vocab">`site:`</font>
    + 1: three to five year old from disadvantaged inner city area
    + 2: four year old from advantaged suburban area
    + 3: from advantaged rural area
    + 4: from disadvantaged rural area
    + 5: from Spanish speaking home


```{r}
# read in dataset
sesame <- read_csv("data/sesame.csv")
```

```{r}
# mean-center relevant continuous variables, make categorical variables factors
sesame <- sesame %>% 
  mutate(viewcat = as.factor(viewcat), 
         site = as.factor(site), 
         prenumbCent = prenumb - mean(prenumb), 
         preletCent = prelet - mean(prelet), 
         ageCent = age - mean(age),
         viewenc = ifelse(viewenc == 1, "Encouraged", "Not Encouraged"))
```
<br>

### Questions

1. We will build a model to predict how often a child in this study watched *Sesame Street*. What type of model should we build? Why? 

2. Describe how you would conduct exploratory data analysis. What plots and/or summary statistics would you include? What information would you learn from the exploratory data analysis?

```{r warning = F, message = F}
model1 <- multinom(viewcat ~ site + viewenc + prenumbCent + preletCent + ageCent,
                   data = sesame)
kable(tidy(model1, conf.int=TRUE, exponentiate = FALSE),
      format = "markdown")
```
<br>

3. Interpret the intercept associated with the odds of `viewcat == 2` versus `viewcat == 1`. 

4. Interpret the effect of the numbers pretest score on the odds of viewership. 


5. The primary objective of the experiment was to understand the effect of encouragement `viewenc` on viewership. Does encouragement have a significant effect on viewership? If so, describe the effect. Otherwise, explain why not. 

6.  We want to test if there are any significant interactions with `viewenc` and the pretests. We create a model that includes the variables from `model1` along with `viewenc*preletCent` and `viewenc*prenumbCent`. 

```{r echo = T, warning = F, message = F}
model2 <- multinom(viewcat ~ site + viewenc + prenumbCent + preletCent + ageCent + 
                     viewenc*preletCent + viewenc*prenumbCent,
                   data = sesame)
```

The results from the drop-in-deviance test are shown below. Is there evidence of a significant interaction effect? Explain. 

```{r}
anova(model1, model2, test = "Chisq")
```

7. How would you assess the appropriateness of the model flit? Describe the plots, tables, and/or calculations you would create to assess model fit. 

### References

Data from [http://www2.stat.duke.edu/~jerry/sta210/sesamelab.html](http://www2.stat.duke.edu/~jerry/sta210/sesamelab.html)

